---
title: "proStat"
author: "Logan Lewis"
date: '2022-04-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pmartR)
library(tidyverse)
library(STRINGdb)
library(stringr)
```

```{r}
proStat <- readRDS("/Users/lewi052/MAP/STRINGdb_exploration/proStat2.RDS")
colnames(proStat)
```

```{r}
length(proStat$Protein)
length(unique(proStat$Protein))
length(unique(proStat$Protein_Identifier))
```


#data exploration

```{r}
attributes(proStat)
```

```{r}
plot(proStat)
```
^ these are the 6 groups well be looking at

## Adding a column of protein identifiers to feed into STRINGdb

```{r}
tmp <- str_split(proStat$Protein, "\\W")[[1]][3]
Gene_Name <- c(Gene_Name, str_split(tmp, "_")[[1]][1])
Gene_Name <- c()
for(i in 1:nrow(proStat)){
  tmp <- str_split(proStat$Protein, "\\W")[[i]][3]
  Gene_Name <- c(Gene_Name, str_split(tmp, "_")[[1]][1])
}
proStat <- mutate(proStat, Gene_Name = Gene_Name)
saveRDS(proStat, file = "proStat2.RDS")
```

## Filtering by Up/Down regulation across condition comparisons

```{r}
Up_HvS <- filter(proStat, `Flag_A_Healthy Control_vs_Severe` == "1")
Up_MvH <- filter(proStat, `Flag_A_Mild_vs_Healthy Control` == "1")
Up_MvS <- filter(proStat, `Flag_A_Mild_vs_Severe` == "1")
Down_HvS <- filter(proStat, `Flag_A_Healthy Control_vs_Severe` == "-1")
Down_MvH <- filter(proStat, `Flag_A_Mild_vs_Healthy Control` == "-1")
Down_MvS <- filter(proStat, `Flag_A_Mild_vs_Severe` == "-1")
```

```{r}
colnames(proStat)[which(names(proStat) == "P_value_A_Healthy Control_vs_Severe")] <- "P_value"
colnames(proStat)[which(names(proStat) == "Fold_change_Healthy Control_vs_Severe")] <- "Fold_change"
colnames(proStat)[which(names(proStat) == "Flag_A_Healthy Control_vs_Severe")] <- "Flag"

string_db <- STRINGdb$new(version="11.5", species=9606, score_threshold=400, input_directory="/Users/lewi052/MAP/STRINGdb_exploration/STRINGdb_cache")

filtered_stat <- filter(proStat, Flag != 0)
length(unique(filtered_stat$Protein_Identifier))
mapped_stat <- string_db$map(filtered_stat, "Protein_Identifier", removeUnmappedRows = TRUE, takeFirst = F)
length(unique(mapped_stat$Protein_Identifier))
test <- mapped_stat[duplicated(mapped_stat$Protein_Identifier),]

mapped_stat_pval05 <- string_db$add_diff_exp_color( subset(mapped_stat, P_value<0.05), logFcColStr="Fold_change" )
mapped_stat_pval05 <- mutate(mapped_stat_pval05, Fold_change_abs = abs(mapped_stat_pval05$Fold_change))
mapped_stat_pval05 <- mapped_stat_pval05[order(mapped_stat_pval05$`Fold_change_abs`, decreasing = T), ]

payload_id <- string_db$post_payload( mapped_stat_pval05$STRING_id, colors=mapped_stat_pval05$color )

hits <- mapped_stat_pval05[1:150, ]
length(unique(hits$Protein_Identifier))


# png("/Users/lewi052/MAP/network.png", width = 1400, height = 1100)
# string_db$plot_network(hits$STRING_id, payload_id = payload_id)
# dev.off()
```

```{r}
postFormSmart <- function(uri, ..., .params = list(), .opts = curlOptions(url = uri),
                          curl = getCurlHandle(), style = 'HTTPPOST',
                          .encoding = integer(), binary = NA, .checkParams = TRUE,
                          .contentEncodeFun = curlEscape){

    res = postForm(uri, ..., .params = .params, .opts = .opts,
             curl = curl, style = style,
             .encoding = .encoding, binary = binary, .checkParams = .checkParams,
             .contentEncodeFun = .contentEncodeFun)
  
    
    suppressWarnings( if(grepl("The document has moved", res)){
    
      begin <- regexpr("href",res)+6
      mys2=substr(res, begin, 10000000)
      end <- regexpr('"',mys2)-1
      uriNew = substr(mys2, 1, end)
      
      res=postForm(uriNew, ..., .params = .params, .opts = .opts,
               curl = curl, style = style,
               .encoding = .encoding, binary = binary, .checkParams = .checkParams,
               .contentEncodeFun = .contentEncodeFun)
    } )
    
    return(res)
  
}
```


```{r}
#converting given column names to known ones
colnames(proStat)[which(names(proStat) == "P_value_A_Healthy Control_vs_Severe")] <- "P_value"
colnames(proStat)[which(names(proStat) == "Fold_change_Healthy Control_vs_Severe")] <- "Fold_change"
colnames(proStat)[which(names(proStat) == "Flag_A_Healthy Control_vs_Severe")] <- "Flag"

#mapping based on UNIPROT identifier
string_db <- STRINGdb$new(version="11.5", species=9606, score_threshold=400, input_directory="/Users/lewi052/MAP/STRINGdb_exploration/STRINGdb_cache")
mapped_stat <- string_db$map(proStat, "Protein_Identifier", removeUnmappedRows = TRUE )

#sorting by absolute value of logFC
mapped_stat <- mutate(mapped_stat, Fold_change_abs = abs(mapped_stat$Fold_change))
mapped_stat <- mapped_stat[order(mapped_stat$`Fold_change_abs`, decreasing = T), ]

nodes <- mapped_stat[1:150, c("Protein_Identifier", "STRING_id", "Gene_Name")]

identifiers = ""
for (id in nodes$Protein_Identifier) {
    identifiers = paste(identifiers, id, sep = "%0d")
}
urlStr = "https://string-db.org/api/tsv/network?"
params <- list(identifiers = identifiers, species = 9606, required_score = 400)
inters <- read_tsv(postFormSmart(urlStr, .params = params))
```


## Creating STRINGdb object

```{r}
string_db <- STRINGdb$new(version="11", species=9606, score_threshold=400, input_directory="/Users/lewi052/MAP/STRINGdb_exploration/STRINGdb_cache")
```

```{r}
Up_HvS <- filter(Up_HvS, `P_value_A_Healthy Control_vs_Severe` < 0.05)
hist(Up_HvS$`Fold_change_Healthy Control_vs_Severe`, breaks = 20)
```


## Mapping proteins using STRINGdb

```{r}
Up_HvS <- filter(Up_HvS, `Fold_change_Healthy Control_vs_Severe` > 0.15)
Up_HvS <- Up_HvS[order(Up_HvS$`Fold_change_Healthy Control_vs_Severe`, decreasing = T), ]
example1_mapped <- string_db$map(Up_HvS, "Protein_Identifier", removeUnmappedRows = TRUE )
```
## protein network (top 200 proteins, sorted by p-value)

```{r}
hits <- example1_mapped$STRING_id[1:150]
string_db$plot_network(hits)
```
## Enrichment and annotations

```{r}
enrichment <- string_db$get_enrichment( hits )
head(enrichment, n=20)
```
```{r}
 annotations <- string_db$get_annotations( hits )
head(annotations, n=20)
```

```{r}
colnames(proStat)[which(names(proStat) == "P_value_A_Healthy Control_vs_Severe")] <- "P_value"
colnames(proStat)[which(names(proStat) == "Fold_change_Healthy Control_vs_Severe")] <- "Fold_change"
colnames(proStat)[which(names(proStat) == "Flag_A_Healthy Control_vs_Severe")] <- "Flag"
colnames(proStat)[which(names(proStat) == "Count_Healthy Control")] <- "Count"
colnames(proStat)[which(names(proStat) == "Mean_Healthy Control")] <- "Mean"

selection <- proStat %>% select(Protein, Count, Mean, P_value, Fold_change, Flag, Protein_Identifier)

filtered_stat <- filter(selection, Flag == 1) %>% filter(P_value < 0.05)

filtered_stat <- filtered_stat[order(filtered_stat$`Fold_change`, decreasing = T), ]

mapped_stat <- string_db$map(filtered_stat, "Protein_Identifier", removeUnmappedRows = TRUE )
hits_test <- mapped_stat$STRING_id[1:150]

enrichment_test <- string_db$get_enrichment(hits_test)
annotations_test <- string_db$get_annotations(hits_test)
```

## Clustering

```{r}
clustersList <- string_db$get_clusters(mapped_stat_pval05$STRING_id[1:150])
```

```{r}
clusters <- c()
for (x in mapped_stat_pval05$STRING_id[1:150]) {
  clusters <- c(clusters, grep(x, clustersList))
}
for (y in clusters){
  if (sum(clusters==y) == 1){
    clusters[which(clusters == y)] <- "No Cluster"
  }
}
```


```{r}
par(mfrow=c(2,2))
for(i in seq(1:4)){
string_db$plot_network(clustersList[[i]])
}
```

```{r}
Down_HvS <- filter(Down_HvS, `P_value_A_Healthy Control_vs_Severe` < 0.05)
hist(Down_HvS$`Fold_change_Healthy Control_vs_Severe`, breaks = 20)
```


## Mapping proteins using STRINGdb

```{r}
Down_HvS <- filter(Down_HvS, `Fold_change_Healthy Control_vs_Severe` < -0.65)
#Down_HvS <- Down_HvS[order(Down_HvS$`Fold_change_Healthy Control_vs_Severe`), ]
example2_mapped <- string_db$map(Down_HvS, "Protein_Identifier", removeUnmappedRows = TRUE )
```
## protein network (top 200 proteins, sorted by p-value)

```{r}
hits <- example2_mapped$STRING_id
string_db$plot_network(hits)
```
```{r}
enrichment_down <- string_db$get_enrichment( hits )
head(enrichment_down, n=20)
```

```{r}
 annotations_down <- string_db$get_annotations( hits )
head(annotations_down, n=20)
```

```{r}
clustersList_down <- string_db$get_clusters(example1_mapped$STRING_id)
```

```{r}
par(mfrow=c(2,2))
for(i in seq(1:4)){
string_db$plot_network(clustersList_down[[i]])
}
```